### 2017-04-04 paper/oopsla17-submission/learning.tex
some (possibly) better text for logistic regression

\subsection{Learning Algorithms}
\label{sec:models}
\ES{TODO: mention that this is a SUPERVISED, CLASSIFICATION problem}
%
There are many learning algorithms to choose from, existing
on a spectrum that balances expressiveness with ease of training (and of
interpreting the learned model).
%
In this section we present the learning algorithms we will investigate,
beginning with the simplest.
%
For each algorithm, we describe \ES{TODO...}

\paragraph{Logistic regression}
The simplest model we investigate is logistic regression, which learns a
linear function of the features.
%
The goal is to learn a set of weights $W$ and biases $b$ such that the
function
%
$$
y = Wx + b
$$
%
effectively maps feature vectors $x$ to labels $y$.

\ES{TODO: talk about logistic regression and the loss function?}

Linear models, like logistic regression, are popular for their ease of
% FIXME: Are you sure you want to call LOGISTIC regression a LINEAR model? 
training and of interpreting the resulting model; however, they can be
limited in their applicability.
%
As the name implies, linear models may perform poorly when asked to
learn an inherently \emph{nonlinear} function.
%
This limitation can be mitigated to some extent by adding more features
to the model (\eg combinations of the existing features \ES{CITE?}), but
the fundamental issue remains.

@@

>>>>

\subsection{Learning Algorithms}
\label{sec:models}
\lstDeleteShortInline{|} % sigh...

Recall that we formulate the localization problem as a supervised
classification problem.
%
This means that we assume the existence of a training set
$D = \{\langle \mathbf{x}_i, y_i \rangle\}_{i=1}^{n}$
of sample inputs $\mathbf{x}$ and their corresponding labels $y$.
%
The learning task is to model the probability $Pr(y\ |\ \mathbf{x})$
of a label being assigned to a given input.

There are many learning algorithms to choose from, existing
on a spectrum that balances expressiveness with ease of training (and of
interpreting the learned model).
%
In this section we present four standard learning algorithms that we
will investigate:
%
(1) logistic regression,
(2) decision trees,
(3) random forests, and
(4) multi-layer perceptrons.
%
For each algorithm, we describe the form of the model it learns, the
training procedure, and any hyper-parameters that must be supplied by a
human.

\paragraph{Logistic regression}
% The simplest model we investigate is logistic regression, which learns a
% linear function of the features.
% %
% The goal is to learn a set of weights $W$ and biases $b$ such that the
% function
% %
% $$
% Pr(y\ |\ \mathbf{x}) = \frac{1}{1 + \mathsf{exp}(-W\mathbf{x} - b)}
% $$
% %
% effectively maps feature vectors $\mathbf{x}$ to labels $y$.

% \ES{TODO: talk about logistic regression and the loss function?}

The simplest model we investigate is logistic regression, which can be
understood in two steps.
%
First, we compute the \emph{evidence} for a label $y$ as a linear
function of the input $\mathbf{x}$, with a set of weights $w_i$ and a
bias $b$ that will be learned from the training set.
$$
\mathsf{ev} = \sum_i w_i \mathbf{x}_i + b
$$
Then we use the logistic function $f(x) = \frac{1}{1 + e^{-x}}$ to
compress the evidence term into the $[0,1]$ interval, which can then be
interpreted as the probability that the label $y$ should be applied.
%
This gives us the final model
$$
Pr(y\ |\ \mathbf{x}) = \frac{1}{1 + \mathsf{exp}(-\sum_i w_i \mathbf{x}_i - b)}
$$

Training a logistic regression entails finding optimal values for the
weights and bias, such that a \emph{cost} function is minimized.
%
In practice, the cost function used for logistic regression is the
\emph{cross entropy}, a measure of the similarity between two
probability distributions, between the predicted labels and the ground
truth.
%
The search for optimal values is typically done via \emph{stochastic
  gradient descent}, an iterative method for optimizing a cost function.
%
Starting with an initial estimate of the weights (\eg a normal
distribution), one repeatedly makes predictions for the training data,
computes the cost function for the predictions, and updates the
parameters according to the gradient of the cost function.
%
This process is repeated until the parameters converge, or until a time
limit is exhausted.
%
Stochastic gradient descent requires a hyper-parameter $\eta$, often
called the \emph{learning rate}, which controls how much the weights and
bias are updated at each iteration.

It has been observed \ES{CITE??} that large weights often coincide with
\emph{overfitting} of the model --- where the model performs well on the
training samples but not on the testing samples.
%
Thus, it is common to add a \emph{regularization} term to the cost
function, a popular choice is the L2 norm of the weights $\sum_i w_i^2$,
which has the effect of penalizing the model for learning large weights.
The contribution of the regularization term to the over cost function is
controlled by another hyper-parameter $\lambda$, often called the
regularization rate.

Logistic regression is popular for its ease of training and of
interpreting the resulting model; however, it can be limited in its
applicability.
%
As a generalized linear model, it may perform poorly when asked to
learn an inherently \emph{nonlinear} function.
%
This limitation can be mitigated to some extent by adding more features
to the model (\eg combinations of the existing features \ES{CITE?}), but
the fundamental issue remains.

@@

% ES: keep this at the bottom... sigh
\lstMakeShortInline{|}
